---
title: "Functional Programming with Purrr"
subtitle: "<br/>#brug"
author: "Saurav Ghosh"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

# What is Functional Programming

## *Wikipedia*

“ -a style of building the structure and elements of computer programs - that treats computation as the evaluation of **mathematical functions** and avoids **changing-state** and **mutable data**. It is a **declarative programming** paradigm, which means programming is done with expressions or declarations instead of statements.”


---

# A brief detour- Lambda calculus

$\lambda x.x$

Where the name after λ is the argument and the expression after the (.) is the body of the function. In programming languages, you can rewrite the above expression as:

```{r}
F<-function (x) { 
    return(x) 
}
```


### Example:

\begin{align*}
f(x) =& x^2
\end{align*}

Which can be rewritten as:

\begin{align*}
x =&>x^2
\end{align*}

---

# Lambda contd…

### Example 1
\begin{align*}
p =&\big[1..100\big].filter\big(\big(value\big)\Rightarrow\big\{return\ value\%2==0\big\}\big)
\end{align*}

### Example 2
\begin{align*}
a =&\big[1..50\big] \\
a.map\big(\big(value\big)\Rightarrow\big\{return\ value*2\big\}\big)
\end{align*}

---

# Question

$\lambda x(\lambda y.x+y)$

--
\begin{align*}
const\ F =&x\Rightarrow\big(y\Rightarrow\big(x+y\big)\big)
\end{align*}

--
\begin{align*}
F\big(5\big)\big(10\big)
\end{align*}

---

# Summary

* immutability (thread-safe)
* explicit state management
* side effect programming through data transformation
* expressions vs statements
* higher level functions (function that takes data and a function as arguments to transform the data)

---

# Purrr

Two types of vectors:
* Atomic
* List

---

# Functions

```{r message=FALSE}
library(tidyverse)
mt<-mtcars
mt %>% str
```

---

# Normalize

## Let us normalize displacement and horsepower
### Let us use min-max normalization

```{r}
mt$new_disp<-(mt$disp-min(mt$disp))/(max(mt$disp)-min(mt$disp))
mt$new_hp<-(mt$hp-min(mt$hp))/(max(mt$hp)-min(mt$hp))
summary(mt$new_hp)
summary(mt$new_disp)
```
---

# Write functions

## Let us rewrite the normalization step as a function

```{r}
func_norm<-function(x){
  (x-min(x))/(max(x)-min(x))
}
my_vec<-rnorm(15)
my_vec
# normalized myvec
func_norm(my_vec)
```

---

# Put it all together

## Let us apply the function in the dataset mt

```{r results="hide"}
mt<-mt %>% mutate(norm_disp=func_norm(mt$disp)) 
mt<-mt %>% mutate(norm_hp=func_norm(mt$hp))
```

```{r}
summary(mt$norm_disp)
summary(mt$norm_hp)
```

---

# Use for loop

```{r}
mt_for<-mtcars
for(i in seq_along(mtcars)){
  mt_for[i]<-func_norm(mtcars[[i]])
}
summary(mt_for)
```

---

# Can we do better?

## Using map
```{r}
mt %>% 
  map(function(x){
    func_norm(x)
  })
```
---

# Even better

## Use anonymous function
```{r}
mt %>% 
  map(~func_norm(.))
```
---

# Return as data frame

```{r}
mt %>% 
  map_df(~func_norm(.))
```

---
# purrr::map(.x,.f,...)

*map* iterates over a list and returns a list.

* .x list (or vector) to iterate over
* .f function to apply over that list
* ... things that get passed from map() to .f

```{r}
my_list=list(a=1:10,b=20:30)
map(my_list,~mean(.))
```
---

# more maps

* map *list*
* map_lgl *logical*
* map_int *integer*
* map_dbl *double*
* map_chr *character*

```{r}
# map_chr
map_chr(my_list,~mean(.))

# map_dbl
map_dbl(my_list,~mean(.))

# map_lgl
# map_lgl(my_list,~mean(.))
# Error: Can't coerce element 1 from a double to a logical
```

---

# map2

## For two lists use map2

`map2(.x, .y, .f, ...)`

```{r}
a<-c(1,3,5,7,9)
b<-c(2,4,6,8,10)
# map2
map2(.x=a,.y = b,~sum(.x,.y))
```
---

# pmap

## For more, use pmap

`pmap(.l, .f, ...)`
```{r}
a<-c(1,2,3,4)
b<-c(5,6,7,8)
c<-c(4,3,2,1)
d<-c(8,7,6,5)
pmap(list(a,b,c,d),~sum(a,b,c,d))
```

---

# Many models with purrr

## Gapminder data

```{r message=FALSE}
library(gapminder)
library(broom)
gapminder %>%
  group_by(country) %>%
  nest() %>%  
  mutate(fit = map(data, ~ lm(lifeExp ~ year, data = .x)))
```

---

# Many models contd...

## View model parameters using broom package

```{r}
gp<-gapminder %>%
  group_by(country) %>%
  nest() %>%  
  mutate(fit = map(data, ~ lm(lifeExp ~ year, data = .x))) %>% 
  mutate(tidied=map(fit,tidy)) %>% 
  mutate(glanced=map(fit,glance)) %>% 
  mutate(augmented=map(fit,augment))
gp
```
---

# R-squared results

```{r}
gapminder %>%
  group_by(country) %>%
  nest() %>%  
  mutate(fit = map(data, ~ lm(lifeExp ~ year, data = .x))) %>% 
  mutate(tidied=map(fit,tidy)) %>% 
  mutate(glanced=map(fit,glance)) %>% 
  mutate(augmented=map(fit,augment)) %>% 
  mutate(rsq=map(glanced,~.[["r.squared"]])) %>% 
  unnest(rsq) %>% 
  arrange(desc(rsq)) %>% 
  top_n(5)
```


---

# View Tidy results

## Unnest the variable tidied

```{r}
unnest(gp,tidied)
```
---
# View Glance results

## Unnest the variable glanced

```{r}
unnest(gp,glanced)
```
---
# View Augmented results

## Unnest the variable augmented

```{r}
unnest(gp,augmented)
```

---

# Other map functions

* keep
* discard
* map_if
* every
* some

More details on [Hooked on Data blog by Emily Robinson](https://hookedondata.org/going-off-the-map/).

---

# Thank you!

## QnA
